MADA Data Analysis Project

### 2021-11-28 code/analysis_code/04-Mods.P2.RandomForest.rmd
Post Holiday Push
Push of analysis completion. Loosing internet 11/28 and will update coherent manuscript edits by Monday Due date.

Bonus- try to figure out how to change into 3 levels?
```{r}
P2.Plus.Cat<-
  P2%>%
  mutate(
# if greater than ct 38, this is a negative detection, thus less 38 is a positive in detection
    Ct.Cat = ifelse(Ct >=38,"No","Yes"), 
# There are some NAs, they are undetectable, change only in the Ct.Cat to No as a factor    
    Ct.Cat=ifelse(is.na(Ct.Cat), "No",Ct.Cat),
# These are factors so mutate the chr to fctrs
    Ct.Cat=factor(Ct.Cat)
    )%>%
#Remove/adjust so there are no character based factors
    select(PCR, RH,TEMP, Method, Tool, Input, Detector, Ct, Ct.Cat)%>%
  na.omit()

@@

#bootstrapping- random sample of data with replacment
bootstraps(P2.Plus.Cat, v=5)
```
Preprocessing
Using parsnims to create a recipe for fitting the model
No standardization
adding dummy variables are added
```{r}
fit_recipe <- 
# set a recipe
  recipe(Ct ~ ., data = P2.Plus.Cat) %>%
# code all categorical variables as dummy variables
  step_dummy(all_nominal())

print(fit_recipe)
```
Null Model
Ct is a continuions outcome of numbers-- need to use RMSE as performance metric.
```{r}
RMSE_null_train <- sqrt(sum( (train_data$Ct - mean(train_data$Ct))^2 )/nrow(train_data))

@@

```{r}
#workflow
rf_wf <- workflow() %>%
  add_model(rf_model) %>% 
  add_recipe(fit_recipe)
```

@@

# tune the model, optimizing RMSE
rf_tune_res <- rf_wf %>%
  tune_grid(
            resamples = cv_data, #CV object
            grid = rf_grid, # grid of values to try
            metrics = metric_set(rmse) 
  )
```

```{r}
#see a plot of performance for different tuning parameters
rf_tune_res %>% autoplot()
```
```{r}
# get the tuned model that performs best 
best_rf <- rf_tune_res %>%  select_best(metric = "rmse")

@@

# fitting best performing model
best_rf_fit <- best_rf_wf %>% 
  fit(data = train_data)
rf_pred <- predict(best_rf_fit, train_data)
```

```{r}
#pull out the fit object
x <- best_rf_fit$fit$fit$fit
#plot variable importance
vip::vip(x, num_features = 20)
```
Meaning that the point of Tool B is the most important/predicitive variable for Ct outcome

```{r}
#predicted versus observed
plot(rf_pred$.pred,train_data$Ct, xlim =c(10,50), ylim=c(10,50))
abline(a=0,b=1, col = 'red') #45 degree line, along which the results should fall
```
```{r}
#residuals
plot(rf_pred$.pred-train_data$Ct)
abline(a=0,b=0, col = 'red') #straight line, along which the results should fall
```
```{r}
rf_perfomance <- rf_tune_res %>% show_best(n = 5)
print(rf_perfomance)
```

>>>>

Bonus- try to figure out how to change into 3 levels?
```{r}
P2.Plus.Cat<-
  P2%>%
  mutate(
# if greater than ct 38, this is a negative detection, thus less 38 is a positive in detection
    Ct.Cat = ifelse(Ct >=38,"No","Yes"), 
# There are some NAs, they are undetectable, change only in the Ct.Cat to No as a factor    
    Ct.Cat=ifelse(is.na(Ct.Cat), "No",Ct.Cat),
# These are factors so mutate the chr to fctrs
    Ct.Cat=factor(Ct.Cat)
    )%>%
#Remove/adjust so there are no character based factors
    select(RH,TEMP, Method, Tool, Input, Detector, Ct, Ct.Cat)%>%
  na.omit()

@@

```
Preprocessing
Using parsnims to create a recipe for fitting the model
No standardization
adding dummy variables are added
```{r}
fit_recipe <- 
# set a recipe
  recipe(Ct ~ ., data = P2.Plus.Cat) %>%
# code all categorical variables as dummy variables
  step_dummy(all_nominal())

print(fit_recipe)
```
Null Model
Ct is a continuations outcome of numbers-- need to use RMSE as performance metric.
```{r}
RMSE_null_train <- sqrt(sum( (train_data$Ct - mean(train_data$Ct))^2 )/nrow(train_data))

@@

```{r}
#workflow
rf_wf <- workflow() %>%
  add_model(rf_model) %>% 
  add_recipe(fit_recipe)
```
Model tuning with a grid.

@@

# tune the model, optimizing RMSE
rf_tune_res <- rf_wf %>%
  tune_grid(
            resamples = cv_data, #CV object
            grid = rf_grid, # grid of values to try
            metrics = metric_set(rmse) 
  )
```
Plot the performance of the grid tune results.
```{r}
#see a plot of performance for different tuning parameters
rf_tune_res %>% autoplot()
```

```{r}
# get the tuned model that performs best 
best_rf <- rf_tune_res %>%  select_best(metric = "rmse")

@@

# fitting best performing model
best_rf_fit <- best_rf_wf %>% 
  fit(data = train_data)
rf_pred <- predict(best_rf_fit, train_data)
```
Use vip to learn about variable importance in the modeling
```{r}
#pull out the fit object
x <- best_rf_fit$fit$fit$fit
#plot variable importance
P2RFBestFit<-vip::vip(x, geom="point",aesthetics = list(color = "blue", shape = 17, size = 4))

print(P2RFBestFit)
```
Meaning that the point of Tool B is the most important/predictive variable for Ct outcome, makes sense as the top 3 are tool related and all have different "concentrations".
Also not pressurizing that Dilution and Input are important.
Method has virtually 0 importance in the Ct value during the modeling. 
```{r warning=FALSE}
#save figure
figure_file = here("results","P2-RF-BestFit.png")
ggsave(filename = figure_file, plot=P2RFBestFit) 
```

```{r}
#predicted versus observed
plot(rf_pred$.pred,train_data$Ct )
abline(a=0,b=1, col = 'red') #45 degree line, along which the results should fall
```
```{r}
#residuals
plot(rf_pred$.pred-train_data$Ct)
abline(a=0,b=0, col = 'red') #straight line, along which the results should fall
```
```{r}
rf_perfomance <- rf_tune_res %>% show_best(n = 5)
print(rf_perfomance)
```

### 2021-11-28 code/analysis_code/05-Mods.P2.TargetedML.rmd
Post Holiday Push
Push of analysis completion. Loosing internet 11/28 and will update coherent manuscript edits by Monday Due date.

```{r}
P2.minus1<-
  P2%>%
  filter(Tool!="B")%>% #remove B because this is actually a -3 dilution. Looking specifically at -1, keep SD range. 
  mutate(
# if greater than ct 38, this is a negative detection, thus less 38 is a positive in detection
    Ct.Cat = ifelse(Ct >=38,"No","Yes"), 
# There are some NAs, they are undetectable, change only in the Ct.Cat to No as a factor    
    Ct.Cat=ifelse(is.na(Ct.Cat), "No",Ct.Cat),
# These are factors so mutate the chr to fctrs
    Ct.Cat=factor(Ct.Cat)
    )%>%
#Remove/adjust so there are no character based factors
    select(PCR, RH,TEMP, Method, Tool, Input, Detector, Ct, Ct.Cat,Dilution)%>%
  na.omit()

@@

```{r}
# Put 3/4 of the data into the training set; this leaves 1/4 of the data to be used to test 
data_split <- initial_split(P2.minus1, prop = 3/4)

## Create data frames for the two sets:
train_data <- training(data_split)
test_data  <- testing(data_split)
```

```{r}
#View the train and test data
glimpse(train_data)
glimpse(test_data)
```

@@

2.1. Many engines Examples
```{r}
# fit linear model
lm_lm    <- lm(Ct ~ ., data = P2.minus1)
# fit general linear model
lm_glm   <- glm(Ct ~ ., data = P2.minus1, 
                family = gaussian)
#########
#<atempt to use later>
# meta engine (aggregator) method=<method name>; see parsnips
#lm_caret <- train(Ct ~ ., data = P2.Plus.Cat, 
#                 method = "glm")
```

2.4. Resampling - CV
```{r }
# create CV object from training data
cv_data <- rsample::vfold_cv(train_data, v = 5, repeats = 5, strata = 'Ct')

print(cv_data)

############## below are similar functions but need to try

#apply CV in ML algorithm to process the ML model to each resample
vfold_cv(P2.minus1, v=5)

#bootstrapping- random sample of data with replacment
bootstraps(P2.minus1, v=5)
```
Preprocessing
Using parsnips to create a recipe for fitting the model
No standardization
adding dummy variables are added
```{r}
fit_recipe <- 
# set a recipe
  recipe(Ct ~., data = P2.minus1) %>%
# code all categorical variables as dummy variables
  step_dummy(all_nominal())

print(fit_recipe)
```
Null Model
Ct is a continuions outcome of numbers-- need to use RMSE as performance metric.
```{r}
RMSE_null_train <- sqrt(sum( (train_data$Ct - mean(train_data$Ct))^2 )/nrow(train_data))

RMSE_null_test <- sqrt(sum( (test_data$Ct - mean(test_data$Ct))^2 )/nrow(test_data))

print(RMSE_null_train)

#predict(lm_lm, train_data=Ct)
```

```{r}
rf_model <- rand_forest() %>%
  set_args(mtry = tune(),     
    trees = tune(),
    min_n = tune()
  ) %>%
  # select the engine/package that underlies the model
  set_engine("ranger",
             num.threads = 18, #for some reason for RF, we need to set this in the engine too
             importance = "permutation") %>%
  # choose either the continuous regression or binary classification mode
  set_mode("regression")   
```

```{r}
#workflow
rf_wf <- workflow() %>%
  add_model(rf_model) %>% 
  add_recipe(fit_recipe)
```

```{r}
#tuning grid
rf_grid  <- expand.grid(mtry = c(3, 4, 5, 6), min_n = c(40,50,60), trees = c(500,1000)  )

# tune the model, optimizing RMSE
rf_tune_res <- rf_wf %>%
  tune_grid(
            resamples = cv_data, #CV object
            grid = rf_grid, # grid of values to try
            metrics = metric_set(rmse) 
  )
```

```{r}
#see a plot of performance for different tuning parameters
rf_tune_res %>% autoplot()
```
```{r}
# get the tuned model that performs best 
best_rf <- rf_tune_res %>%  select_best(metric = "rmse")

# finalize workflow with best model
best_rf_wf <- rf_wf %>% finalize_workflow(best_rf)

# fitting best performing model
best_rf_fit <- best_rf_wf %>% 
  fit(data = train_data)
rf_pred <- predict(best_rf_fit, train_data)
```

```{r}
#pull out the fit object
x <- best_rf_fit$fit$fit$fit
#plot variable importance
P2RFBestFitminus1<-vip::vip(x, num_features = 20)
print(P2RFBestFitminus1)
```
Interesting so Dilution is obviously more important followed by Input. It's more surprizing that the Method is one of the least important in this comparison. Infact the temperature and relative humidity are more impractically thatn the method.
```{r warning=FALSE}
#save figure
figure_file = here("results","P2-RF-BestFit-minus1.png")
ggsave(filename = figure_file, plot=P2RFBestFitminus1) 
```

```{r}
#predicted versus observed
plot(rf_pred$.pred,train_data$Ct, xlim =c(10,50), ylim=c(10,50))
abline(a=0,b=1, col = 'red') #45 degree line, along which the results should fall
```
```{r}
#residuals
plot(rf_pred$.pred-train_data$Ct)
abline(a=0,b=0, col = 'red') #straight line, along which the results should fall
```
```{r}
rf_perfomance <- rf_tune_res %>% show_best(n = 5)
print(rf_perfomance)
```

@@

Tree model
```{r}
tree_model <-  decision_tree() %>% 
  set_args( cost_complexity = tune(), tree_depth = tune(), min_n = tune()) %>% 
  set_engine("rpart") %>% 
  set_mode("regression")      
```
Workflow
```{r}
#set workflow
tree_wf <- workflow() %>%
  add_model(tree_model) %>%
  add_recipe(fit_recipe)
```
Tree Model
```{r}
#tuning grid
tree_grid <- dials::grid_regular(cost_complexity(), tree_depth(), min_n(), levels = 5)
#tune the model
tree_tune_res <- tree_wf %>% 
  tune::tune_grid(
    resamples = cv_data,
    grid = tree_grid,
    metrics = yardstick::metric_set(rmse) 
  )
```

```{r}
#see a plot of performance for different tuning parameters
tree_tune_res %>% autoplot()
```
```{r}
# get the tuned model that performs best 
best_tree <- tree_tune_res %>%  select_best(metric = "rmse")
# finalize workflow with best model
best_tree_wf <- tree_wf %>% finalize_workflow(best_tree)
# fitting best performing model
best_tree_fit <- best_tree_wf %>% 
                 fit(data = train_data)
#predicting outcomes for final model
tree_pred <- predict(best_tree_fit, train_data)
```

```{r}
rpart.plot(extract_fit_parsnip(best_tree_fit)$fit)
```

>>>>

```{r}
P2.minus1<-
  P2%>%
  filter(Tool!="B")%>% #remove B because this is actually a -3 dilution. Looking specifically at -1, keep SD range.
  mutate(
# if greater than ct 38, this is a negative detection, thus less 38 is a positive in detection
    Ct.Cat = ifelse(Ct >=38,"No","Yes"), 
# There are some NAs, they are undetectable, change only in the Ct.Cat to No as a factor    
    Ct.Cat=ifelse(is.na(Ct.Cat), "No",Ct.Cat),
# These are factors so mutate the chr to fctrs
    Ct.Cat=factor(Ct.Cat)
    )%>%
#Remove/adjust so there are no character based factors
    select(RH,TEMP, Method, Tool, Input, Detector, Ct, Ct.Cat,Dilution)%>%
  na.omit()

@@

```{r}
# Put 3/4 of the data into the training set; this leaves 1/4 of the data to be used to test 
data_split2 <- initial_split(P2.minus1, prop = 3/4)

## Create data frames for the two sets:
train_data2 <- training(data_split2)
test_data2  <- testing(data_split2)
```

```{r}
#View the train and test data
glimpse(train_data2)
glimpse(test_data2)

table(train_data2$Method)%>%
  prop.table()
table(test_data2$Method)%>%
  prop.table()
```

@@

2.1. Many engines Examples
```{r}
# fit linear model
lm_lm    <- lm(Ct ~ ., data = P2.minus1)
# fit general linear model
lm_glm   <- glm(Ct ~ ., data = P2.minus1, 
                family = gaussian)
#########
#<atempt to use later>
# meta engine (aggregator) method=<method name>; see parsnips
lm_caret <- train(Ct ~ Method + Dilution + Input, data = P2.minus1, 
                 method = "glm")
```

2.4. Resampling - CV
```{r }
# create bootstrap resample object from training data
cv_data_boots <- bootstraps(train_data2, v=5)
print(cv_data_boots)
```
Preprocessing
Using parsnips to create a recipe for fitting the model
No standardization
adding dummy variables are added
```{r}
fit_recipe2 <- 
# set a recipe specific to the Dilution and Input
  recipe(Ct ~Dilution+Input+Method, data = P2.minus1) %>%
# code all categorical variables as dummy variables
  step_dummy(all_nominal())

print(fit_recipe2)
```
Null Model
Ct is a continuions outcome of numbers-- need to use RMSE as performance metric.
```{r}
RMSE_null_train2 <- sqrt(sum( (train_data2$Ct - mean(train_data2$Ct))^2 )/nrow(train_data2))

RMSE_null_test2 <- sqrt(sum( (test_data2$Ct - mean(test_data2$Ct))^2 )/nrow(test_data2))

print(RMSE_null_train2)

predict(lm_glm, train_data=Ct)
```

```{r}
rf_model2 <- rand_forest() %>%
  set_args(mtry = tune(),     
    trees = tune(),
    min_n = tune()
  ) %>%
  # select the engine/package that underlies the model
  set_engine("ranger",
             num.threads = 18, #for some reason for RF, we need to set this in the engine too
             importance = "permutation") %>%
  # choose either the continuous regression or binary classification mode
  set_mode("regression")   
```

```{r}
#workflow
rf_wf2 <- workflow() %>%
  add_model(rf_model2) %>% 
  add_recipe(fit_recipe2)
```
Try bootstrapping in CV workflow
```{r}
#tuning grid
rf_grid2  <- expand.grid(mtry = c(3, 4, 5, 6), min_n = c(40,50,60), trees = c(500,1000)  )

# tune the model, optimizing RMSE
rf_tune_res2 <- rf_wf2 %>%
  tune_grid(
            resamples = cv_data_boots, #CV object
            grid = rf_grid2, # grid of values to try
            metrics = metric_set(rmse) 
  )
```

```{r}
#see a plot of performance for different tuning parameters
rf_tune_res2 %>% autoplot()
```
```{r}
# get the tuned model that performs best 
best_rf2 <- rf_tune_res2 %>%  select_best(metric = "rmse")

# finalize workflow with best model
best_rf_wf2 <- rf_wf2 %>% finalize_workflow(best_rf2)

# fitting best performing model
best_rf_fit2 <- best_rf_wf2 %>% 
  fit(data = train_data2)
rf_pred2 <- predict(best_rf_fit2, train_data2)
```

```{r}
#pull out the fit object
x2 <- best_rf_fit2$fit$fit$fit
#plot variable importance
P2RFBestFitminus1<-vip::vip(x2, num_features = 10, geom="point", aesthetics = list(color = "red", shape = 17, size = 4))
print(P2RFBestFitminus1)
```
Interesting so Dilution is obviously more important followed by Input. It's more surprizing that the Method is one of the least important in this comparison. Infact the temperature and relative humidity are more impractically thatn the method.
```{r warning=FALSE}
#save figure
figure_file = here("results","P2-RF-BestFit-minus1.png")
ggsave(filename = figure_file, plot=P2RFBestFitminus1) 
```

```{r}
#predicted versus observed
plot(rf_pred2$.pred,train_data2$Ct)
abline(a=0,b=1, col = 'red') #45 degree line, along which the results should fall
```
```{r}
#residuals
plot(rf_pred2$.pred-train_data2$Ct)
abline(a=0,b=0, col = 'red') #straight line, along which the results should fall
```
```{r}
rf_perfomance2 <- rf_tune_res2 %>% show_best(n = 5)
print(rf_perfomance2)
```

@@

Tree model
```{r}
tree_model2 <-  decision_tree() %>% 
  set_args( cost_complexity = tune(), tree_depth = tune(), min_n = tune()) %>% 
  set_engine("rpart") %>% 
  set_mode("regression")      
```
Workflow
```{r}
#set workflow
tree_wf2 <- workflow() %>%
  add_model(tree_model2) %>%
  add_recipe(fit_recipe2)
```
Tree Model
```{r}
#tuning grid
tree_grid2 <- dials::grid_regular(cost_complexity(), tree_depth(), min_n(), levels = 5)
#tune the model
tree_tune_res2 <- tree_wf2 %>% 
  tune::tune_grid(
    resamples = cv_data_boots,
    grid = tree_grid2,
    metrics = yardstick::metric_set(rmse) 
  )
```

```{r}
#see a plot of performance for different tuning parameters
tree_tune_res2 %>% autoplot()
```
```{r}
# get the tuned model that performs best 
best_tree2 <- tree_tune_res2 %>%  select_best(metric = "rmse")
# finalize workflow with best model
best_tree_wf2 <- tree_wf2 %>% finalize_workflow(best_tree2)
# fitting best performing model
best_tree_fit2 <- best_tree_wf2 %>% 
                 fit(data = train_data2)
#predicting outcomes for final model
tree_pred2 <- predict(best_tree_fit2, train_data2)
```

```{r}
minusTree<-
  rpart.plot(extract_fit_parsnip(best_tree_fit2)$fit,type = 5, tweak = 1.2)

print(minusTree)
```
Each node shows: predicted values, and percentage of observations in node.
So if the dilution is still the lead in the effects of the Ct. - Makes sense as that's the spiking/manipulating level 
If the dilution is greater than -1.5, effects of the input factor strongly in the Ct recovery, with a little input from the method if the input is less than 650uL

### 2021-11-28 products/manuscript/Manuscript.Rmd
Post Holiday Push
Push of analysis completion. Loosing internet 11/28 and will update coherent manuscript edits by Monday Due date.

---
title: "MADA Data Analysis Project"
subtitle: ""
author: Monica Chan
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
documentclass: article
output:
  bookdown::html_document2: 
    toc: false
  bookdown::word_document2: 
    toc: false
    number_sections: true
bibliography: ../references.bib
csl: ../apa.csl
---

@@

Applying MADA learned techniques to a RT-PCR data set to determine the effectiveness of sample enhancement strategies for environmental surface sampling of low bio-burden viruses.

@@

The global Coronavirus Disease 2019 (COVID-19) pandemic continues to persist due to the timing of an available vaccine, varying adherence to COVID-19 public health safety recommendations, and the emergence of SARS-CoV-2 (SC2) variants. Environmental reservoirs are a reported, indirect transmission route for viral pathogens [@Marcenac2021-sz]. Although many viruses have a low infectious dose, they can be difficult to detect due to the low bio-burden in the environment. The ability to detect viral pathogens is critical to identifying hidden reservoirs which can inform infection control [@Paton2021-jc]. Culture-based methods can be labor and time intensive. The use and limitations of commercial "target capture" (TC) technologies are varied compared to molecular approaches utilizing automated extraction. The objective of this study is improve outbreak response through rapid pathogen detection by evaluating a TC technology's ability to capture SC2 from complex matrices.

### Questions/Hypotheses to be addressed

@@

Using established standard methods for environmental sampling with a self-contained polyurethane swab from stainless steel surface, a TC approach (NTP; Nanotrap® Magnetic Virus Particles, Ceres Nanosciences) was compared to an automated extraction method (MX48; Maxwell® RSC Viral Total Nucleic Acid Purification Kit, Promega) for recovery of heat-inactivated SC2 from the polyurethane swab as detected by rRT-PCR [@Wittwer2012-vt][@Shah2021-mb].

@@

```{r}
knitr::include_graphics("../../results/P2-SD-Plot.png")
```

@@

```{r}
knitr::include_graphics("../../results/ASK-P2P1-Tools.png")
```

@@

#### Phase 2 Comparing Method Serial Dilutions (CDC vs NTP) with standard input of 300uL

```{r}
knitr::include_graphics("../../results/ASK-P2P2-Methods.png")
```

@@

```{r}
knitr::include_graphics("../../results/ASK-P2P3-Methods0300.png")
```

@@

```{r}
knitr::include_graphics("../../results/ASK-P2P4-MethodsALL.png")
```

@@

```{r}
knitr::include_graphics("../../results/ASK-P2P5-NTP-SD.png")
```

@@

```{r}
knitr::include_graphics("../../results/ASK-P2P6-NTP.png")
```

@@

### Data Modeling

@@

Figure \@ref(fig:resultfigure) shows a scatterplot figure produced by one of the R scripts.

@@

## Strengths and Limitations

Need more replicates, but this is a time consuming processes just to get the data.

Additional materials need to be considered [@Jansson2020-cg].

## Conclusions

>>>>

---
title: "MADA Data Analysis Project"
subtitle: ""
author: Monica Chan
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
documentclass: article
output:
  bookdown::html_document2: 
    toc: false
  bookdown::word_document2: 
    toc: false
    number_sections: true
bibliography: ../references.bib
csl: ../PLoS.csl
---

@@

The global Coronavirus Disease 2019 (COVID-19) pandemic continues to persist due to the timing of an available vaccine, varying adherence to COVID-19 public health safety recommendations, and the emergence of SARS-CoV-2 (SC2) variants. Environmental reservoirs are a reported, indirect transmission route for viral pathogens. Although many viruses have a low infectious dose, they can be difficult to detect due to the low-bioburden in the environment. The ability to detect viral pathogens is critical to identifying hidden reservoirs which can inform infection control. Culture-based methods can be labor and time intensive. The use and limitations of commercial “target capture” (TC) technologies are varied compared to molecular approaches utilizing automated extraction. The objective of this study is improved outbreak response through rapid pathogen detection by evaluating a TC technology’s ability to capture SC2 from complex matrices.

@@

The global emergence of novel human coronavirus, SARS-CoV-2 (SC2) in 2019 became a global health concern. SC2 causes severe respiratory tract infections in humans with still unknown future complications. With the virus' high infection rate and rapid spreading capabilities via human-to-human transmission

The global Coronavirus Disease 2019 (COVID-19) pandemic continues to persist due to the timing of an available vaccine, varying adherence to COVID-19 public health safety recommendations, and the emergence of SARS-CoV-2 (SC2) variants. Environmental reservoirs are a reported, indirect transmission route for viral pathogens [@Tanner2021;@Marcenac2021-sz]. Although many viruses have a low infectious dose, they can be difficult to detect due to the low bio-burden in the environment. The ability to detect viral pathogens is critical to identifying hidden reservoirs which can inform infection control [@Paton2021-jc]. Culture-based methods can be labor and time intensive. The use and limitations of commercial "target capture" (TC) technologies are varied compared to molecular approaches utilizing automated extraction. The objective of this study is improve outbreak response through rapid pathogen detection by evaluating a TC technology's ability to capture SC2 from complex matrices.

The virus is capable of human-to-human transmission via inhalation of infected aerosols. Previous evaluations of viral fragment longevity have tested negative and indicate the possibility of no present virus or below assay limit of detection [@Parker2020].
### Questions/Hypotheses to be addressed

@@

Using established standard methods for environmental sampling with a self-contained polyurethane swab from stainless steel surface, a TC approach (NTP; Nanotrap® Magnetic Virus Particles, Ceres Nanosciences) was compared to an automated extraction method (MX48; Maxwell® RSC Viral Total Nucleic Acid Purification Kit, Promega) for recovery of heat-inactivated SC2 from the polyurethane swab as detected by rRT-PCR [@Wittwer2012-vt;@Shah2021-mb;@Barclay2020].

@@

```{r P2-SD, fig.cap="Serial Dilution Comparisons within Phase 2 data", echo=FALSE}
knitr::include_graphics("../../results/P2-SD-Plot.png")
```

@@

```{r P2-Tool-AB, echo=FALSE, fig.cap="Phase 2 Comparison of A and B tool recovery"}
knitr::include_graphics("../../results/ASK-P2P1-Tools.png")
```

@@

#### Phase 2 Comparing Method Serial Dilutions (CDC vs NTP) with standard input of 300uL
MOVE TO SUPPLEMENT
```{r}
knitr::include_graphics("../../results/ASK-P2P2-Methods.png")
```

@@

```{r P2Tools-Method, fig.cap="P2 Direct Comparisons of Recovery by Methods used", echo=FALSE}
knitr::include_graphics("../../results/ASK-P2P3-Methods0300.png")
```

Data was limited to Phase 2 data from this point on wards to limit variability in the surface recovery

@@

@@

@@

```{r echo=FALSE}
knitr::include_graphics("../../results/ASK-P2P6-NTP.png")
```

@@

### Data Modeling
```{r echo=FALSE}
knitr::include_graphics("../../results/P2-RF-BestFit.png")
```

```{r ImportanceFactor, fig.cap="Bootstrapped Importance Factor excluding Tools ", echo=FALSE}
knitr::include_graphics("../../results/P2-RF-BestFit-minus1.png")
```

```{r minusDecisionTree, fig.cap="Bootstrapped Decision Tree of Limited Factors.", echo=FALSE}
knitr::include_graphics("../../results/P2-DecisionTree-minus1.png")
```

@@

Figure \@ref(fig:minusDecisionTree) shows a bootstrapped decision tree figure.

@@

## Strengths and Limitations
RT-PCR does not distinguish between infectious and non-infectious virus. Additional work will need to be completed to better compare the recovery and estimate the true risk of SC2 in the environment.
Need more replicates, but this is a time consuming processes just to get the data.

Additional materials need to be considered [@Jansson2020-cg].
Previous studies have demonstrated capture techniques can overcome limitations of conventional studies, and if applied to environmental samples can yield better data to enlighten the field [@Gasc2018].
## Conclusions
The use of nanoparticles does result in the recovery of Ct values in the assay.
with the results of the analysis of data there is no significant difference between the automated extraction method and the nanoparticles.
Using some ML modeling we can see that there are some recorded variables that affect the Ct values- Dilution of the sample and Input.
